---
title: "SVARtca: Transmission Channel Analysis for Structural VAR Models in R"
abstract: >
  We introduce SVARtca, an R package implementing Transmission Channel Analysis
  (TCA) for structural VAR models following Wegner, Lieb, and Smeekes (2025).
  TCA decomposes impulse response functions into contributions from distinct
  causal transmission channels using a systems form representation and directed
  acyclic graph (DAG) path analysis. The package supports three decomposition
  modes: overlapping channels, exhaustive 3-way decomposition, and exhaustive
  4-way decomposition via the inclusion-exclusion principle. SVARtca provides a
  high-level interface that works directly with vars::VAR objects, a low-level
  systems form interface for custom SVAR specifications, and publication-ready
  ggplot2 visualizations. It is a parallel R implementation of the original
  MATLAB toolbox by Wegner, Lieb, and Smeekes, making this methodology
  accessible to the large R-based macroeconometrics community. The package is
  open-source under the GPL-3 license.
author:
  - name: Muhammad Abdullah Alkhalaf
    affiliation: Rufyq Elngeh for Academic and Business Services
    address: Riyadh, Saudi Arabia
    orcid: 0009-0002-2677-9246
    email: muhammedalkhalaf@gmail.com
type: package
output:
  rjtools::rjournal_article:
    self_contained: yes
    toc: no
bibliography: paper.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE, comment = "#>")
```

# Introduction

Structural Vector Autoregression (SVAR) models are the workhorse of empirical
macroeconomics for tracing the dynamic effects of structural shocks. Impulse
response functions (IRFs) quantify the total effect of a shock on each variable
over time, but they do not reveal *through which channels* the effects propagate.

\citet{Wegner2025} introduced Transmission Channel Analysis (TCA), a methodology
that decomposes IRFs into channel-specific contributions using the systems form
representation and directed acyclic graph (DAG) path analysis. The original
implementation is a MATLAB toolbox, limiting accessibility for the large R-based
macroeconometrics community. The `vars` package \citep{Pfaff2008vars} provides
SVAR estimation and IRF computation but not channel decomposition.

`SVARtca` bridges this gap with a native R implementation that interfaces directly
with `vars::VAR` objects while also providing a low-level systems form interface
for custom SVAR specifications.

# Methodology

## Systems form representation

Consider a K-variable VAR(p) model:

$$\mathbf{y}_t = \sum_{j=1}^{p} \mathbf{\Phi}_j \mathbf{y}_{t-j} + \mathbf{u}_t$$

With structural identification $\mathbf{u}_t = \mathbf{B} \boldsymbol{\varepsilon}_t$,
the structural moving average representation is:

$$\mathbf{y}_t = \sum_{s=0}^{\infty} \mathbf{\Theta}_s \boldsymbol{\varepsilon}_{t-s}$$

where $\mathbf{\Theta}_s = \mathbf{\Psi}_s \mathbf{B}$ and $\mathbf{\Psi}_s$ are
the reduced-form MA coefficients.

## Transmission channel decomposition

TCA converts the VAR to a systems form where the contemporaneous impact matrix
$\mathbf{B}$ defines a DAG over the variables. A transmission channel from shock
variable $i$ to target variable $j$ through intermediate variable $m$ is defined
by the set of causal paths that pass through $m$.

The channel-specific contribution is computed by zeroing out all paths in the
systems form that do not pass through the specified intermediate, then
recomputing the IRF from this modified system.

## Decomposition modes

Three modes are available:

- **Overlapping:** Computes per-channel contributions independently. When channels
  share intermediate variables, contributions may sum to more than the total IRF.
- **Exhaustive 3-way:** Decomposes the total IRF into three non-overlapping
  components: the inclusive channel (through at least one specified intermediate),
  the exclusive channel (through none of the specified intermediates), and the
  direct effect.
- **Exhaustive 4-way:** Applies the inclusion-exclusion principle to decompose
  into four components when two intermediates are specified: each channel alone,
  the joint channel (through both), and the direct effect.

# Usage

## High-level interface with vars::VAR

The `tca_from_var` function provides the simplest workflow:

```{r svartca-var-example}
library(SVARtca)
library(vars)

# Simulate a 4-variable macroeconomic system
set.seed(42)
n <- 200
interest_rate <- arima.sim(list(ar = 0.8), n = n)
credit <- 0.5 * interest_rate + arima.sim(list(ar = 0.6), n = n)
exchange_rate <- -0.3 * interest_rate + 0.2 * credit +
                 arima.sim(list(ar = 0.7), n = n)
output <- 0.4 * credit - 0.2 * exchange_rate +
          arima.sim(list(ar = 0.5), n = n)

macro_data <- ts(data.frame(
  interest_rate = interest_rate,
  credit = credit,
  exchange_rate = exchange_rate,
  output = output
))

# Estimate VAR
var_est <- VAR(macro_data, p = 2, type = "const")

# Transmission channel analysis
result <- tca_from_var(var_est,
                       from = "interest_rate",
                       intermediates = c("credit", "exchange_rate"),
                       h = 20, mode = "overlapping")

# Print results for output variable
print(result, target = 3)

# Publication-ready plot
plot_tca(result, target = "output")
```

## Low-level systems form interface

For custom SVAR specifications (e.g., sign restrictions, external instruments),
construct the systems form directly:

```{r svartca-lowlevel}
# Example with custom identification
K <- 4
p <- 2

# Suppose we have custom Phi0 (contemporaneous) and As (lag matrices)
# from an externally identified SVAR
Phi0 <- diag(K)  # placeholder - use actual identification
As <- array(0, dim = c(K, K, p))  # placeholder - use estimated coefficients

sf <- tca_systems_form(Phi0, As, h = 20, order = 1:K)
result <- tca_analyze(from = 1, B = sf$B, Omega = sf$Omega,
                      intermediates = c(2, 3), K = K, h = 20,
                      order = 1:K, mode = "exhaustive_4way",
                      var_names = c("interest_rate", "credit",
                                    "exchange_rate", "output"))
```

## Visualization

The `plot_tca` function produces `ggplot2` objects showing the total IRF alongside
channel-specific contributions:

```{r svartca-plot}
# Overlapping channels plot
p1 <- plot_tca(result, target = "output")

# Customize with ggplot2
library(ggplot2)
p1 + theme_minimal() +
  labs(title = "Monetary Policy Transmission to Output")
```

# Comparison with existing implementations

The `vars` package \citep{Pfaff2008vars} provides comprehensive SVAR estimation
including Cholesky, Blanchard--Quah, and structural identification, along with
IRF computation and forecast error variance decomposition (FEVD). However, it
does not decompose IRFs into channel-specific contributions.

The `svars` package extends `vars` with data-driven identification methods
(statistical identification via independent components, heteroskedasticity-based
identification) but similarly provides only total IRFs without channel
decomposition.

`SVARtca` is complementary to both packages: it takes as input either a
`vars::VAR` object or custom coefficient matrices and adds the TCA decomposition
layer. The key methodological contribution is the ability to answer questions
like "How much of the effect of a monetary policy shock on output operates
through the credit channel versus the exchange rate channel?" --- questions that
standard IRF analysis cannot address.

The original MATLAB implementation by \citet{Wegner2025} provides the same
methodology. `SVARtca` offers equivalent functionality in R with additional
convenience features: direct `vars::VAR` integration, `ggplot2`-based
visualization, and tidy output objects compatible with the R ecosystem.

# Summary

`SVARtca` makes Transmission Channel Analysis accessible to R users, enabling
decomposition of SVAR impulse responses into channel-specific contributions.
The package provides both high-level and low-level interfaces, supports three
decomposition modes, and produces publication-ready visualizations. It is
available on CRAN under GPL-3.

\bibliography{paper}
